---
title: "Team Standings Report"
author: Nishat Mohammad
date: 01/19/2024
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

## Report 1 on Team Standings
```{r Q7_Team_Standing}
# Pull code from my .R file I made earlier for this project
source("DA5030.A6.108.v1.Nishat.Mohammad.R", echo=TRUE)

### Question 7. ####
#Create an R Notebook and build a “report” that lists the teams and then produces the “standings”, i.e., a table that lists the teams in order of points. Award 2 points to the winner, 1 for a team that loses in overtime, and 0 for a team that lost in regulation.
#### Answers ####

# Create a df for the table 
points_table <- as.data.frame(matrix(ncol= 5, nrow=0))
colnames(points_table) <- c("game_num", "winning_team", "Winner_point", "losing_team","losser_point")

my_teams_df
# Get the points in th points table
for (row in 1: nrow(my_teams_df)){
  points_table[row,1] <- row
  if (my_teams_df$ot[row]==FALSE){
      if (my_teams_df$home.team.goals[row] > my_teams_df$away.team.goals[row]){
        points_table[row,2] <- my_teams_df$home.team[row]
        points_table[row,3] <- 2
        points_table[row,4] <- my_teams_df$away.team[row]
        points_table[row,5] <- 0
      }else {
        points_table[row,2] <- my_teams_df$away.team[row]
        points_table[row,3] <- 2
        points_table[row,4] <- my_teams_df$home.team[row]
        points_table[row,5] <- 0
      }
  } else{
    if (my_teams_df$home.team.goals[row] > my_teams_df$away.team.goals[row]){
        points_table[row,2] <- my_teams_df$home.team[row]
        points_table[row,3] <- 2
        points_table[row,4] <- my_teams_df$away.team[row]
        points_table[row,5] <- 1
      }else {
        points_table[row,2] <- my_teams_df$away.team[row]
        points_table[row,3] <- 2
        points_table[row,4] <- my_teams_df$home.team[row]
        points_table[row,5] <- 1
      }
    } 
}
points_table
# Get the team names
team_names <- unique(c(unique(points_table$winning_team), unique(points_table$losing_team)))

# Create the Standings table
standings_table <- as.data.frame(matrix(ncol= 2, nrow=length(team_names)))
colnames(standings_table) <- c("team_name", "total_points")

# Add the team names column
standings_table[1:(length(team_names)),1]<- team_names

for (row in 1:nrow(standings_table)){
  total_win_point<-points_table$Winner_point[points_table$winning_team==standings_table$team_name[row]]
  total_loss_point<- points_table$losser_point[points_table$losing_team==standings_table$team_name[row]]
  total_point <- sum(total_loss_point,total_win_point)
  standings_table$total_points[row] <- total_point
}
standings_table
tot_points<- standings_table$total_points

```
### Team List
The Team List is:  
`r paste0('"',hm_team,'"')`

### Standings
The standings for the respective teams are:  
`r paste(format(round(tot_points,0), big.mark = ",", scientific = FALSE))`  
View all details in the tables above.  


## Report 2 on Team Standings
```{r Q8_Report2}
### Question 8. ####
#Add another section to the R Notebook that creates a different table based on the winner getting 3 points, the overtime winner gets 2 points, and the loser in overtime gets 1 point.
#### Answers ####
# Create a df for the second table 
points_table2 <- as.data.frame(matrix(ncol= 5, nrow=0))
colnames(points_table2) <- c("game_num", "winning_team", "Winner_point", "losing_team","losser_point")

# Get the points in th points table2
for (row in 1: (length(my_teams_df$home.team.goals))){
  points_table2[row,1] <- row
  if (my_teams_df$ot[row]==FALSE){
      if (my_teams_df$home.team.goals[row] > my_teams_df$away.team.goals[row]){
        points_table2[row,2] <- my_teams_df$home.team[row]
        
        points_table2[row,3] <- 3
        points_table2[row,4] <- my_teams_df$away.team[row]
        points_table2[row,5] <- 0
      }else {
        points_table2[row,2] <- my_teams_df$away.team[row]
        points_table2[row,3] <- 3
        points_table2[row,4] <- my_teams_df$home.team[row]
        points_table2[row,5] <- 0
      }
  } else{
    if (my_teams_df$home.team.goals[row] > my_teams_df$away.team.goals[row]){
        points_table2[row,2] <- my_teams_df$home.team[row]
        points_table2[row,3] <- 2
        points_table2[row,4] <- my_teams_df$away.team[row]
        points_table2[row,5] <- 1
      }else {
        points_table2[row,2] <- my_teams_df$away.team[row]
        points_table2[row,3] <- 2
        points_table2[row,4] <- my_teams_df$home.team[row]
        points_table2[row,5] <- 1
      }
    } 
}
points_table2

# Get the team names
team_names2 <- unique(c(unique(points_table2$winning_team), unique(points_table2$losing_team)))

# Create the Standings table
standings_table2 <- as.data.frame(matrix(ncol= 2, nrow=length(team_names2)))
colnames(standings_table2) <- c("team_name", "total_points")

# Add the team names column
standings_table2[1:(length(team_names2)),1]<- team_names2

# Get the total score and add to the second standings table
for (row in 1:nrow(standings_table2)){
  total_win_point2<-points_table2$Winner_point[points_table2$winning_team==standings_table2$team_name[row]]
  total_loss_point2<- points_table2$losser_point[points_table2$losing_team==standings_table2$team_name[row]]
  total_point2 <- sum(total_loss_point2,total_win_point2)
  standings_table2$total_points[row] <- total_point2
}
standings_table2
tot_points2 <- standings_table2$total_points

```
### Team List
The team list remains the same as in report 1:  
`r paste0('"',hm_team,'"')`  


### Standings  
The Standings have now changed based of winner is regular time getting 3 points and 0 for loosers, winners in over time get 2 points while loosers in over time get 1. Please view the standings for the respective teams:  

`r paste(format(round(tot_points2,0), big.mark = ",", scientific = FALSE))`  

View all details in the tables above.  

## Analysis of Standings Report 1 and 2
```{r Q9_Report analysis}
# Bar plots for the reports
# Report 1
colors <- c("red", "orange", "yellow", "green")
pnts <- c(standings_table$total_points)
name_teams <- c(standings_table$team_name)

# Create a bar plot 
barplot(height = pnts, width = 2, space = 1, names.arg = name_teams,
        main = "Bar Plot of Standings Report 1",
        xlab = "Teams",
        ylab = "Points",
        col = colors,
        border = "blue",
        ylim = c(0, 20) )


# Report 2
pnts2 <- c(standings_table2$total_points)
name_teams2 <- c(standings_table2$team_name)

# Create a bar plot
barplot(height = pnts2, width = 2, space = 1, names.arg = name_teams2,
        main = "Bar Plot of Standings Report 2",
        xlab = "Teams",
        ylab = "Points",
        col = colors,
        border = "blue",
        ylim = c(0, 20) )
```

### Bar plots for Standings
The Bar plots above visualize the two reports with Arctic Avalanche having the highest points and Glacial Guardians having the lowest points in both reports. Over all both reports are supportive of each others findings. 

The points scored by the teams in Report 1 are `r pnts` for `r name_teams` respectively.  
The points scored by the teams in Report 2 are `r pnts2` for `r name_teams2` respectively.  

